---
title: "Tame Impala"
description: |
  I dive into Tame Impala's music through the spotifyr package and Spotify's API.
author:
  - name: Joe DeCesaro
    url: {}
date: 12-03-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(spotifyr)
library(tidyverse)
```

```{r, include = FALSE}
#setup
id <- "858051c943024e02a7fc87cbf2c7d1bd"
secret <- "abaa4cc4c64c4b7aa1214753b5cda04d"
SPOTIFY_CLIENT_ID = id
SPOTIFY_CLIENT_SECRET = secret
access_token <- get_spotify_access_token(client_id = Sys.getenv("SPOTIFY_CLIENT_ID"), client_secret = Sys.getenv("SPOTIFY_CLIENT_SECRET"))
authorization <- get_spotify_authorization_code(client_id = Sys.getenv("SPOTIFY_CLIENT_ID"), client_secret = Sys.getenv("SPOTIFY_CLIENT_SECRET"))
```

```{r}
# Some Spotify Constants
tame_id <- "5INjqkS1o8h1imAzPqGZBb"
```


```{r}
tame_artist_catalog_data <- get_artist(tame_id)
tame_album_artist_data <- get_artist_albums(tame_id, 
                                            include_groups = c("album"))
```


```{r}
# Feature data for all songs as of 12/4/2021
tame_feature_data <- get_artist_audio_features(artist = tame_id, 
                                  authorization = get_spotify_access_token(), 
                                  include_groups = c("album"))

# Below does not return artist lyric information for some reason.
# tame_discography <- get_discography(artist = "tame impala", 
#                                   authorization = get_spotify_access_token())
```

```{r}
currents_data <- get_album_tracks("79dL7FLiJFOO0EoehUHQBv", 
                                  authorization = access_token)

the_slow_rush_data <- get_album_tracks("31qVWUdRrlb8thMvts0yYL")
                                       
```

```{r}
# top track data, default is in the US, can call other countries by other ID's
top_tracks <- get_artist_top_tracks(tame_id)
```


```{r}
# Getting my data
my_profile <- get_my_profile(authorization = get_spotify_authorization_code())

my_top_artists <- get_my_top_artists_or_tracks(type = "artists", 
                                               time_range = "long_term",
                                               authorization = authorization)
my_top_tracks <- get_my_top_artists_or_tracks(type = "tracks", 
                                               time_range = "long_term",
                                               authorization = authorization)
```


# Some possible ways to disect music:
* speechiness across albums
* Danceability = “Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. A value of 0.0 is least danceable and 1.0 is most danceable.”
* Valence = “A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry).”
* Energy = “Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. Typically, energetic tracks feel fast, loud, and noisy. For example, death metal has high energy, while a Bach prelude scores low on the scale. Perceptual features contributing to this attribute include dynamic range, perceived loudness, timbre, onset rate, and general entropy.”
* Sentiment analysis

* Compare my most played tame songs to various factors
* Sonic scroe from tayloR post
