---
title: "The ggplot() Function"
description: |
  A short summary of the `ggplot()` function within the tidyverse library.
author:
  - name: Joe DeCesaro
    url: {}
date: 08-10-2021
output:
  distill::distill_article:
    self_contained: false
preview: images/yogi.jpg
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(palmerpenguins)
```

#### The `ggplot()` function:

`ggplot()` allows for the easy plotting of various types of graphic representations of data frames. These plots can start out as simple but can also be highly complex and customized. Below is an example of a simple point graph of data from the Palmer Penguins data set.

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```
Note that it is critical to specify the type of plot desired after the `ggplot()` function. Below is an example of using the

```{r}
ggplot(data = penguins, aes( x = flipper_length_mm, y = body_mass_g)) +
  geom_jitter()
```

## stopped first draft here and will continue to discuss customization of graphs and other things with examples.