---
title: "The Basics of the ggplot() Function"
description: |
  A short introduction to the `ggplot()` function within the tidyverse library.
author:
  - name: Joe DeCesaro
    url: {}
date: 08-10-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: yes
categories:
  - R 
  - tidyverse
preview: images/ggplot_ex.png
---


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(palmerpenguins)
```

#### The basics of the `ggplot()` function:

`ggplot()` allows for the easy plotting of various types of graphic representations of data frames. These plots can start out as simple but can also be highly complex and customized. Below is an example of a simple point graph of data from the Palmer Penguins data set.

```{r, include=TRUE}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```
Note that it is critical to specify the type of plot desired after the `ggplot()` function. Below is an example of using the `geom_line()` function where above uses the `geom_point()` function.

```{r, include=TRUE}
ggplot(data = penguins, aes( x = flipper_length_mm, y = body_mass_g)) +
  geom_line()
```

Looking at the scatter plot above (not the line plot) does not provide much useful information from the data set. In data science it is important to call out specific groups within your data. In the Palmer Penguins data, three separate species of penguins are observed. Let's take a look at how we can separate the data by species within the plot.

```{r, include=TRUE}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species))
```
By seperating the data by species we can now clearly draw some distinct conclusions from the data for each species. For example, we know that Gentoo penguins will generally be the largest penguins.

It is always a good idea to take a look at your data BEFORE you begin working with it. By looking at the `penguins` data prior to plotting it, I would have known that there were three species within this data set and that I should probably have seperated them in the plot from the beginning.

This is a very brief introduction to the `ggplot()` function. Future posts will add to the foundations set in this introduciton.